Make any stupid plot you want

### 2021-02-04 draft/sections/implementation.tex
code data section + refactor continous and discontinous line into one source object

To prototype our model, we implemented the artist classes for the scatter and line plots shown in figure~\ref{fig:code_scatter_line} because they differ in every attribute of our artist: different visual channels \vchannel\ that composite to different marks \vmark\ with different continuities \vindex\.  We make use of the Matplotlib figure and axes artists \cite{hunterArchitectureOpenSource,hunterMatplotlib2DGraphics2007} so that we can initially focus on the data to graphic transformations.

@@

\subsection{Artist Class $\vartist^{\prime}$}
The artist is the piece of the matplotlib architecture that constructs an internal representation of the graphic that the render then uses to draw the graphic. In the prototype artist, \mintinline{python}{transform} is a dictionary of the form \mintinline{python}|{parameter:(variable, encoder)}| where parameter is a component in \vfiber, variable is a component in \dfiber,  and the \vchannel\ encoders are passed in as functions or callable objects. The data bundle \dtotal\ is passed in as a \mintinline{python}{data} object. By binding data and transforms to \vartisteq\ inside \mintinline{python}{__init__}, the \mintinline{python}{draw} method is a fully specified artist \vartist.

@@

    def draw(self, renderer, *args, **kwargs):
        # query data for a vertex table K
        view = self.data.view('vertex') 
        visual = {p: encoder(view.get(f, None)) for 
                     p, (f, encoder) in self.transforms.items()}
        self.assemble(visual)
        # call the renderer that will draw based on properties
        super().draw(renderer, *args, **kwargs)

@@

    def draw(self, renderer, *args, **kwargs):
        # query data source for edge table
        view = self.data.view('edge')
        visual = {p: encoder(view.get(f, None)) for 
                     p, (f, encoder) in self.transforms.items()}
        self.assemble(visual)
        super().draw(renderer, *args, **kwargs)

@@

\subsubsection{Encoders \vchannel}
As mentioned above, the encoding dictionary is specified by the visual fiber component, the corresponding data fiber component, and the mapping function. The visual parameter serves as the dictionary key because the visual representation is constructed from the encoding applied to the data  $\vsection = \vchannel \circ \dsection$. For the scatter plot, the mappings for the visual fiber components $\vfiber=(x,y, facecolors, s)$ are defined as
\begin{minted}{python}
cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
transforms = {'y': ('v2', lambda x: x),       
              'x': ('v4', lambda x: x),
             'facecolors': ('v3', cmap), 
             's':(None ,lambda _: itertools.repeat(.02))}
\end{minted}
where the position \textit{(x,y)} \vchannel\ transformers are identity functions. The size \textit{s} transformer is not acting on a component of \dfiber\, instead it is a \vchannel\ that returns a constant value. While size could be embedded inside the \mintinline{python}{assembly} function, it is added to the transformers to illustrate user configured visual parameters that could either be constant or mapped to a component in \dfiber. The identity and constant \vchannels\ are explicitly implemented here to demonstrate their implicit role in the visual pipeline, but they could be optimized away. More complex encoders can be implemented as callable classes, such as 
\begin{minted}{python}
class Categorical:
    def __init__(self, mapping):
        # check that the conversion is to valid colors
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

The check for equivariance can be implemented trivially such that it is independent of data or encoder.
\begin{minted}{python}
def is_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    return sorted(m1) == sorted(m2)
\end{minted}
In this example, \mintinline{python}{is_nominal} checks for equivariance of permutation group actions by applying the encoder to a set of values, shuffling values, and checking that (value, encoding) pairs remain the same. \note{not sure where/how these checks should be incorporated, maybe as part of transform dict?}

\subsubsection{Data \dtotal}
what, why, how?

\subsubsection{Case Study: Iris}

>>>>

To prototype our model, we implemented the artist classes for the scatter and line plots shown in figure~\ref{fig:code_scatter_line} because they differ in every attribute: different visual channels \vchannel\ that composite to different marks \vmark\ with different continuities \vindex\.  We make use of the Matplotlib figure and axes artists \cite{hunterArchitectureOpenSource,hunterMatplotlib2DGraphics2007} so that we can initially focus on the data to graphic transformations.

@@

\subsection{Artist Class $\vartist^{\prime}$}
\label{sec:code_artist}
The artist is the piece of the matplotlib architecture that constructs an internal representation of the graphic that the render then uses to draw the graphic. In the prototype artist, \mintinline{python}{transform} is a dictionary of the form \mintinline{python}|{parameter:(variable, encoder)}| where parameter is a component in \vfiber, variable is a component in \dfiber,  and the \vchannel\ encoders are passed in as functions or callable objects. The data bundle \dtotal\ is passed in as a \mintinline{python}{data} object. By binding data and transforms to \vartisteq\ inside \mintinline{python}{__init__}, the \mintinline{python}{draw} method is a fully specified artist \vartist.

@@

    def draw(self, renderer, *args, **kwargs):
        # query data for a vertex table K
        view = self.data.view() 
        visual = {p: encoder(view.get(f, None)) for 
                     p, (f, encoder) in self.transforms.items()}
        self.assemble(visual)
        # call the renderer that will draw based on properties
        super().draw(renderer, *args, **kwargs)

@@

    def draw(self, renderer, *args, **kwargs):
        # query data source for edge table
        view = self.data.view()
        visual = {p: encoder(view.get(f, None)) for 
                     p, (f, encoder) in self.transforms.items()}
        self.assemble(visual)
        super().draw(renderer, *args, **kwargs)

@@

\subsection{Encoders \vchannel}
\label{sec:code_channels}
As mentioned above, the encoding dictionary is specified by the visual fiber component, the corresponding data fiber component, and the mapping function. The visual parameter serves as the dictionary key because the visual representation is constructed from the encoding applied to the data  $\vsection = \vchannel \circ \dsection$. For the scatter plot, the mappings for the visual fiber components $\vfiber=(x,y, facecolors, s)$ are defined as
\begin{minted}{python}
cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
transforms = {'y': ('v1', lambda x: x),       
              'x': ('v3', lambda x: x),
             'facecolors': ('v2', cmap), 
             's':(None ,lambda _: itertools.repeat(.02))}
\end{minted}
where the position \textit{(x,y)} \vchannel\ transformers are identity functions. The size \textit{s} transformer is not acting on a component of \dfiber, instead it is a \vchannel\ that returns a constant value. While size could be embedded inside the \mintinline{python}{assembly} function, it is added to the transformers to illustrate user configured visual parameters that could either be constant or mapped to a component in \dfiber. The identity and constant \vchannel\ are explicitly implemented here to demonstrate their implicit role in the visual pipeline, but they could be optimized away. More complex encoders can be implemented as callable classes, such as 
\begin{minted}{python}
class Categorical:
    def __init__(self, mapping):
        # check that the conversion is to valid colors
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

A test for equivariance can be implemented trivially such that it is independent of data or encoder.
\begin{minted}{python}
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
In this example, \mintinline{python}{is_nominal} checks for equivariance of permutation group actions by applying the encoder to a set of values, shuffling values, and checking that (value, encoding) pairs remain the same. This equivariance test can be implemented as part of the artist or encoder, but for minimal overhead, the equivariant it is implemented as part of the library tests.

\subsection{Data \dtotal}
\label{sec:code_data}
The data input into the \mintinline{pythin}{Artist} will often be a wrapper class around an existing data structure, but must meet the following criteria:
\begin{enumerate}
\item specify the fiber components \dfiber\ and connectivity \dbase\
\item have a \mintinline{view} method that returns an atomic object that encapsulates \dsection\
\item the view object must have \mintinline{get} method that returns a fiber component
\end{enumerate}
To support specifying the fiber bundle, we define an optional \mintinline{python}{FiberBundle} class

\begin{minted}{python}
class FiberBundle:
    def __init__(self, base, fiber):
        """
        base:  {'tables': ['vertex', 'edge', 'face']}
        fiber: {'component name': {'type':, 'monoid':,  'range':}}
        """
        self.base = base
        self.fiber = fiber

    def is_section(self, section):
        """checks if a section is from a given fiber bundle:
        are values in F, are keys in K"""
\end{minted}

that asks the user to specify how \dbase is triangulated and the attributes of \dfiber. The \mintinline{python}{assembly} functions expect tables that match the continuity of the graphic; scatter expects a vertex table because it is discontinuous, line expects an edge table because it is 1D continuous. The fiber informs appropriate choice of \vchannel\, therefore it is a dictionary of attributes of the fiber components.  \note{I've basically stripped this out of the artists above so should I just ditch this section?}

To generate the scatter plot in figure~\ref{fig:code_scatter_line}, we fully specify a dataset with random keys and values in a section chosen at random form the corresponding fiber component. The fiberbundle \mintinline{python}{FB} is a class level attribute since all instances of \mintinline{python}{code}{VertexSimplex} come from the same fiberbundle.
\begin{minted}{python}
class VertexSimplex: #maybe change name to something else
    """Fiberbundle is consistent across all sections
    """
    FB = FiberBundle({'tables': ['vertex']},  
            {'v1': {'type': float,'monoid':'interval','range': [0,1]},
             'v2': {'type': str, 'monoid':'nominal', 'range':['true', 'false']},
             'v3': {'type': float, 'monoid':'interval', 'range':[2,3]}})

    def __init__(self, sid = 45, size=1000, max_key=10**10):
        # create random list of keys
    def tau(self, k):
        # e1 is sampled from F1, e2 from F2, etc...
        return (k, (e1, e2, e3, e4))

    def view(self):
        table = defaultdict(list)
        for k in self.keys:
            table['index'] = k
            # on each iteration, add one (name, value) pair per component
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
                table[name].append(value)
        return table
\end{minted}
The view method returns a dictionary where the key is a fiber component name and the value is a list of values in the fiber component. The table is built one call to \mintinline{python}{tau} at a time, guaranteeing that all the fiber component values are over the same \dbasepoint. Table has a \mintinline{python}{get} method as it is a method on Python dictionaries. In contrast, the line in \mintinline{python}{EdgeSimplex} is defined as the functions \mintinline{python}{_color,_xy} on each edge.

\begin{minted}{python}
class EdgeSimplex: 
    # assign a class level FB attribute
    def __init__(self, num_edges=4, num_samples=1000): 
        self.keys = range(num_edge) #edge id
        # distance along edge
        self.distances = np.linspace(0,1, num_samples)
        # half generlized representation of arcs on a circle
        self.angle_samples = np.linspace(0, 2*np.pi, len(self.keys)+1)

    @staticmethod
    def _color(edge):
        colors = ['red','orange', 'green','blue']
        return colors[edge%len(colors)]

    @staticmethod
    def _xy(edge, distances, start=0, end=2*np.pi):
        # start and end are parameterizations b/c really there is 
        angles = (distances *(end-start)) + start
        return np.cos(angles), np.sin(angles)

    def tau(self, k): #will fix location on page on revision
        x, y = self._xy(k, self.distances, 
                        self.angle_samples[k], self.angle_samples[k+1]) 
        color = self._color(k) 
        return (k, (x, y, color))

    def view(self, simplex):
        table = defaultdict(list)
        for k in self.keys:
            table['index'].append(k)
            # (name, value) pair, value is [x0, ..., xn] for x, y
            for (name, value) in zip(self.FB.fiber.keys(), self.tau(k, simplex)[1]):
                table[name].append(value)
\end{minted}
Unlike scatter, the line \mintinline{python}{tau} method returns the functions on the edge evaluated on the interval [0,1]. By default these means each \mintinline{python}{tau} returns a list of 1000 x and y points and the associated color. As with scatter, \mintinline{python}{view} builds a table by calling \mintinline{python}{tau} for each \dbasepoint\. Unlike scatter, the line table is a list where each item contains a list of points. This bookkeeping of which data is on an edge is used by the \mintinline{python}{assembly} functions to bind segments to their visual properties.

\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/linec_1.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/lined_1.png}
    \end{subfigure}
    \caption{Continuous and discontinuous lines as defined by different data models, but generated with the same \vartisteq=\mintinline{python}{artist}}
    \label{fig:circle_step}
\end{figure}

The graphics in figure~\ref{fig:circle_step} are made using the \mintinline{python}{Line} artist and the \mintinline{python}{Graphline} data source

\begin{minted}{python}
class GraphLine:
    def __init__(self, FB, edge_table, vertex_table, num_samples=1000, connect=False):
        #s set args as attributes and generate distance
        if connect: # test connectivity if edges are continuous
            assert edge_table.keys() == self.FB.F.keys()
            assert is_continuous(vertex_table)

    def tau(self, k, simplex='edge'):
        # evaluates functions defined in edge table
        return(k, (self.edges[c][k](self.distances) for c in self.FB.F.keys()))

    def view(self, simplex='edge'):
        """walk the edge_vertex table to return the edge function
        """
        table = defaultdict(list)
        #sort since intervals lie along number line and are ordered pair neighbors
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i, simplex)[1]):
                table[name].append(value)
        return table
\end{minted}
where if told that the data is connected, the data source will check for that connectivity by constructing an adjacency matrix. The multicolored line is a connected graph of edges with each edge function evaluated on 1000 samples 
\begin{minted}{python}
simplex.GraphLine(FB, edge_table, vertex_table, connect=True)
\end{minted}
while the stair chart is discontinuous and only needs to be evaluated at the edges of the interval 
\begin{minted}{python}
simplex.GraphLine(FB, edge_table, vertex_table, num_samples=2, connect=False)
\end{minted}
such that one advantage of this model is it helps differentiate graphics that have different artists from graphics that have the same artist but make different assumptions about the source data.

\subsection{Case Study: Iris}

### 2021-02-04 draft/sections/math.tex
artist + scatter
q hat + filter out glyph
code data section + refactor continous and discontinous line into one source object

\subsection{Artist}
\label{sec:artist}
%%% line things up such that they're an equivariant maps 
%%% make a little clearer what E' is 
The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The artist \vartist\ is a mapping from \dtotal\ padded with data from $\mathcal{J}(\dtotal)$ to a graphic that is a section \gsection\ in  $\Gamma(\gtotal)$
%%% \Gamma is total
%%% H is point wise -
%%% this map means we cna do this pointwise 
%%% explain the arrows, chase through here
%%% new disagram w/ section

@@

\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i})}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $\{f_{0}, \ldots, f_{n}\} \times (\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber \vfiber\ for a visualization function implemented in Matplotlib}
    \label{tab:mpl_visual_variable_fiber}
\end{table}

@@

\subsection{Making the fiber bundle computable}
\label{sec:triangulization}
\begin{figure}[H]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) records to all records are connected to at least 3 others}
    \label{fig:triangle_simplex}
\end{figure}
One way of expressing the connectivity of records in a dataset is to implement \dbase\ as a simplacial complex, which is a set of simplicies such as those shown in figure~\ref{fig:triangle_simplex}. The advantage of triangulization is that it is general enough to work for more complex topology based visualization methods \cite{heineSurveyTopologybasedMethods2016} while also providing a consistent interface of vertices, edges, and faces for \vindex\ to map into. When triangulated, the simplices encode the continuity in the data

>>>>

\subsection{Artist}
\label{sec:artist}
%%somewhere in this I need to add q \circ nu (\xi^{-1}\tau(s)) = rho (s)
%%% line things up such that they're an equivariant maps 
%%% make a little clearer what E' is 
The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The artist \vartist\ is a mapping from \dtotal\ padded with data from $\mathcal{J}(\dtotal)$ to a graphic that is a section \gsection\ in  $\Gamma(\gtotal)$
%%% \Gamma is total
%%% H is point wise -
%%% this map means we can do this pointwise 
%%% explain the arrows, chase through here
%%% new disagram w/ section?
% so you aren't passing around \tau
% thos maps make sense for \tau(k)\in E -> V
% and then you have \mu(k) 
% and then when you are in \xi*V .... well the arrow goes the WONG way so you didn't just pass \mu(k) inot \xi*V
% so if you want to go from \xi*V to H via Q you need to do it over a PARTICULAR s
% thats what I mean by Q (in this diagram) hiding a choice of s. 
% and then IF you pick an s ...... its is just fine. because it returns \rho(s)
% so types match \rho(s) is in D

@@

\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i})}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber \vfiber\ for a visualization function implemented in Matplotlib}
    \label{tab:mpl_visual_variable_fiber}
\end{table}

@@

\subsection{Making the fiber bundle computable}
\label{sec:triangulization}
\begin{figure}[H]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) records to all records are connected to at least 3 others}
    \label{fig:triangle_simplex}
\end{figure}
%%big data
One way of expressing the connectivity of records in a dataset is to implement \dbase\ as a simplacial complex, which is a set of simplicies such as those shown in figure~\ref{fig:triangle_simplex}. The advantage of triangulization is that it is general enough to work for more complex topology based visualization methods \cite{heineSurveyTopologybasedMethods2016} while also providing a consistent interface of vertices, edges, and faces for \vindex\ to map into. When triangulated, the simplices encode the continuity in the data
